\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-09-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
petites modfis + nettoyage d'une section en conflit

\subsubsection{XXX}

\textbf{Il me semble qu'on avait dit qu'on mettait une 3e robustesse. Non? Faire une transition pour la suite.}

\section{Additive costs and transport costs time trends}\label{sec:results_trends}

In this section, we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects. As underlined by \cite{hummels2007}, the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade.
Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased, independently of any change in transport costs \textit{per se}.
As emphasized by \cite{Hummels_1999} and \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
Our own results tend to challenge this view, as we further show.

@@

The estimation carried out in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), which vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
Starting from these values, for each additive and multiplicative cost component (by transport mode) we extract the pure transport cost measure by the mean of a time fixed effect.
Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the unfitted transport costs measures $\{\Gamma^{t/\widetilde{p}, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total cost respectively), and three series for the fitted values $\{\Gamma^{t/\widetilde{p}}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).

>>>>

\subsubsection{Robustness to aggregation}

Current data are available at the 10-digit level for products (HS10).
Theoretically, they should be available from 1989.
But because of the COVID situation, it was only possible to get 1997-1999 and 2001-2019.
Yet, as Table \ref{tab:baseline10} and 	Figure \ref{fig:robustness_to_agregation} show, the results are similar when using 10-digit product and 5-digit products.

\begin{table}[htbp]
	\centering
	\caption{Transport costs estimates: robustess to agregation} \vspace{5mm} \label{tab:baseline10}
	\input{Table_baseline10.tex}
	\begin{tablenotes}
		\tiny
		\item Statistics for 1997-1999 and 2002-2019
		\item Statistics are obtained weighting each observation by its value relative to total trade flows.
		\item \textbf{Model (a): With ad-valorem transport costs only}
		\item \textbf{Model (b): With additive and ad-valorem transport costs}
	\end{tablenotes}
\end{table}

\begin{figure}[htbp]
	\caption{Transport cost time trends: robustness to agregation}
	\label{fig:robustness_to_agregation}
	\begin{center}
		\includegraphics[height=8cm]{scatter_chronology_baseline_baseline10.png}
	\end{center}
\end{figure}

Two main insights emerge up to now. First, total transport costs have strongly reduced between 1974 and 2019. Second, additive costs represent an important share of those declining transport costs, between 30 and 45\%. In the two following sections, we take stock of those results to draw implications for two important debates in economic analysis. In the next section, we shed light on the origins of the decrease of transport costs over time, by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \emph{per se}. Section \ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of alterations over time to multiplicative and additive costs.

\section{Transport costs time trends: the key part of additive costs}\label{sec:results_trends}

In this section, we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects. As underlined by \cite{hummels2007}, the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade.
Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased, independently of any change in transport costs \textit{per se}.
As emphasized by \cite{Hummels_1999, hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
Our own results tend to challenge this view, as we further show.

@@

The estimation carried out in section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), which vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
Starting from these values, for each additive and multiplicative cost component (by transport mode) we extract the pure transport cost measure by the mean of a time fixed effect.
Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the unfitted transport costs measures $\{\Gamma^{t/\widetilde{p}, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total cost respectively), and three series for the fitted values $\{\Gamma^{t/\widetilde{p}}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
